### 2011-09-09 paper.tex
Acknowledgements

The DP recursion is as follows (with $B,C,u,v,w,x$ defined as in \tabref{RuleTypes})
\begin{eqnarray*}
\lefteqn{\m(i,j,k,l,A) =} \\
& &
\sum_{\rho \in \rulesubset{1}{A}} W(\rho) \sum_{m=i}^j \m(m,j,k,l,B) \mtip(i,m,C) \\
& & +
\sum_{\rho \in \rulesubset{2}{A}} W(\rho) \sum_{n=k}^l \m(i,j,k,n,B) \mtip(n,l,C) \\
& & +
\sum_{\rho \in \rulesubset{3}{A}} W(\rho) \sum_{m=i}^j \m(i,m,k,l,B) \mtip(m,j,C) \\
& & +
\sum_{\rho \in \rulesubset{4}{A}} W(\rho) \sum_{n=k}^l \m(i,j,n,l,B) \mtip(k,n,C) \\
& & +
\sum_{\rho \in \rulesubset{5}{A}} W(\rho) \sum_{m=i}^j \sum_{n=k}^l \m(i,m,n,l,B) \m(m,j,k,n,C) \\
& & +
\sum_{\rho \in \rulesubset{6}{A}} W(\rho) \m(i+|u|,j-|v|,k+|w|,l-|x|,B)
 \Delta(i,u) \Delta(j-|v|,v) \Delta(k,w) \Delta(l-|x|,x)
\end{eqnarray*}

@@

@@

Our understanding of grammars applied to bioinformatics has been uplifted in conversations with
Sean Eddy, Mark Steedman, Bonnie Webber, Aravind Joshi, David Searls, Dan Klein, and Michael Souza.

>>>>

The DP recursion is as follows (with $B,C,u,v,w,x$ defined as in \tabref{RuleTypes})
\begin{eqnarray*}
\lefteqn{\m(i,j,k,l,A) =} \\
& &
\sum_{\rho \in \rulesubset{1}{A}} \sum_{m=i}^j W(\rho) \m(m,j,k,l,B) \mtip(i,m,C) \\
& & +
\sum_{\rho \in \rulesubset{2}{A}} \sum_{n=k}^l W(\rho) \m(i,j,k,n,B) \mtip(n,l,C) \\
& & +
\sum_{\rho \in \rulesubset{3}{A}} \sum_{m=i}^j W(\rho) \m(i,m,k,l,B) \mtip(m,j,C) \\
& & +
\sum_{\rho \in \rulesubset{4}{A}} \sum_{n=k}^l W(\rho) \m(i,j,n,l,B) \mtip(k,n,C) \\
& & +
\sum_{\rho \in \rulesubset{5}{A}} \sum_{m=i}^j \sum_{n=k}^l W(\rho) \m(i,m,n,l,B) \m(m,j,k,n,C) \\
& & +
\sum_{\rho \in \rulesubset{6}{A}} W(\rho) \m(i+|u|,j-|v|,k+|w|,l-|x|,B)
 \Delta(i,u) \Delta(j-|v|,v) \Delta(k,w) \Delta(l-|x|,x)
\end{eqnarray*}

@@

\subsubsection{Length distributions}

To model genomic features, it is extremely useful to augment the TAG framework with length distributions;
that is, to allow the rule weight $\weight(\rho)$ to depend on distances
such as $m-i$, $j-m$, $n-k$, $l-n$, $n-m$, $j-i$, $k-j$, $l-k$, $l-i$ and so on.

Note that the TAG framework implicitly allows some length distributions to be constructed, even without such augmentation;
for example, the ``waiting time'' in a particular state is geometrically-distributed, and by arranging a series of such states,
one can obtain other distributions, e.g. a negative binomial distribution.
One can also obtain any distribution over a random variable $x \leq N$ by using $N$ states,
or approximate any distribution to some degree of precision with a finite number of states.
However, it is often much more convenient and efficient simply to allow $\weight$ to sometimes depend directly on the inside sequence length.

@@

Sean Eddy co-organized an excellent workshop at U.Penn, ``Language Modeling of Biological Data'',
at which IH met Mark Steedman, Bonnie Webber and Aravind Joshi,
and learned from them (and others) the secrets of TAGs, finite-state transducers, and mildly context-sensitive grammars.
David Searls has been a long-time proponent of linguistics applied to bioinformatics.
Dan Klein encouraged us to develop our own variations on grammar frameworks.
Michael Souza introduced us to the pumping lemma and the equivalence of TAGs with various other frameworks.
